---
title: "Predicting Gender"
author: "J.Ojeda"
date: "February 12, 2018"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Can we predict the gender if we know the height ?


## The data


The file [pulse.RData](./pulse.RData) has information about different phisical features of students
```{r}
load("pulse.RData")
head(pulse)
```


Some basic exploratory data analysis:


```{r}
summary(pulse)
plot(pulse[,c(6,7,9)],col=pulse$gender)
```


Some nonparametric density estimator:
```{r}
with(pulse, {
  plot(density(heig[gender=="female"]),xlim=c(150,200))
  lines(density(heig[gender=="male"]),col="red")
  }
)
```


## Use Naive Bayes to __fit__ Gender using height

```{r}
library(naivebayes)
genderHeightNB<-naive_bayes(gender~heig,pulse)
genderHeightNB
```

Notice: 

* A priory probabilities are _relative frequency_ of genders:
```{r}
prop.table( table(pulse$gender) )
```

* Mean and sd(standard deviation) table gives the mean and sd of height given the genders (_for each gender_):
```{r}
aggregate(heig~gender,pulse,function(x) c(mean=mean(x),sd=sd(x)))
```

...But indeed is the mean and sd of a fitted gaussian distributions for the height given the genders
```{r}
plot(genderHeightNB)
```


## Use Naive Bayes to __predict__ Gender using height

```{r}
newData<-data.frame(heig=seq(from=150,to=200,by=10))
predict(genderHeightNB,newData)
```

We can also predict _probabilities_:
```{r}
predict(genderHeightNB,newData,type="prob")
```


A bit better:
```{r}
cbind(newData,pred=predict(genderHeightNB,newData))
```

Let us see how many hits (successes) we do have:
```{r}
(a<-xtabs( ~gender+predict(genderHeightNB),pulse ))
```

Which means a `r round(100*sum(diag(a))/sum(a),2)`% of hits and 
`r round(100*(1-sum(diag(a))/sum(a)),2)`% of misses.


## ...but we haven't been very fair !!

We have considered the same data to fit the _model_ and also to check how good it was... and that is not fair: should be simple to predict with the data you have !! 

Lets consider a _random tests set_ of about a third of data and see what happens: 
```{r}
(test<-sample(nrow(pulse),nrow(pulse)/3))
```

First we fit it:
```{r}
genderHeightNBok<-naive_bayes(gender~heig,pulse[-test,])
genderHeightNBok
```

...and we get:
```{r}
(a<-xtabs( ~gender+predict(genderHeightNBok,pulse[test,]),pulse[test,] ))
(1-sum(diag(a))/sum(a))
```

Which means a `r round(100*sum(diag(a))/sum(a),2)`% of hits and 
`r round(100*(1-sum(diag(a))/sum(a)),2)`% of misses.


...Indeed it would be good to repeat with several random tests sets !!!

__Do you dare to do it youself !!??__








